
# Title: Foundation of R Programming
### Author: Md Rabiul Islam
### Date: "`r Sys.Date()`"
### Output: html_document
### Download and Install R and RStudio

- Download R: [CRAN R Project](https://cran.r-project.org/)
- Download RStudio: [RStudio Desktop](https://posit.co/download/rstudio-desktop/)

# Basic Data Types in R

```{r}
num_var <- 10.5  # Numeric
print(num_var)

int_var <- 10L  # Integer
print(int_var)

char_var <- "Hello, R!" # Character (String)
print(char_var)

bool_var <- TRUE # Logical (Boolean)
print(bool_var)
```

# Arithmetic Operators

```{r}
a <- 10
b <- 5
print(a + b)  # Addition
print(a - b)  # Subtraction
print(a * b)  # Multiplication
print(a / b)  # Division
print(a %% b) # Modulus (remainder)
print(a ^ b)  # Exponentiation
```

# Logical Operators

```{r}
x <- TRUE
y <- FALSE
print(x & y)  # AND
print(x | y)  # OR
print(!x)     # NOT
```

# Data Structures

## Vectors (1D data structure)

```{r}
vec <- c(1, 2, 3, 4, 5)
print(vec)
print(vec[1])   # First element
print(vec[2:4]) # Elements from index 2 to 4
```

## Matrix (2D data structure)

```{r}
mat <- matrix(1:12, nrow=4, ncol=3)
print(mat)
```

## Data Frame

```{r}
df <- data.frame(
  Name = c("Alice", "Bob", "Charlie"),
  Age = c(25, 30, 35),
  Score = c(90, 85, 88)
)
print(df)
print(df$Name)  # Access columns
print(df[1,])   # First row
```

## Lists (Flexible heterogeneous data structure)

```{r}
my_list <- list(Name="Alice", Age=25, Scores=c(90, 85, 88))
print(my_list)
print(my_list$Name)  # Access elements
print(my_list$Scores[1])  # First score
```

# Basic Programming Concepts

## Conditional Statements (if, else)

```{r}
x <- 10
if (x > 5) {
  print("x is greater than 5")
} else {
  print("x is 5 or less")
}
```

## For Loop

```{r}
for (i in 1:5) {
  print(paste("Iteration:", i))
}
```

## While Loop

```{r}
count <- 1
while (count <= 5) {
  print(paste("Count:", count))
  count <- count + 1
}
```

## Defining Functions

```{r}
add_numbers <- function(a, b) {
  return(a + b)
}
print(add_numbers(5, 3))
```

# Data Manipulation

```{r}
#install.packages("dplyr")  # Install dplyr package
library(dplyr)              # Load the package

df <- data.frame(
  Name = c("Alice", "Bob", "Charlie"),
  Age = c(25, 30, 35),
  Score = c(90, 85, 88)
)

# Select columns
df_selected <- select(df, Name, Score)
print(df_selected)

# Filter rows
df_filtered <- filter(df, Age > 25)
print(df_filtered)

# Arrange rows
df_sorted <- arrange(df, desc(Score))
print(df_sorted)

# Add new column
df_mutated <- mutate(df, Pass = Score >= 85)
print(df_mutated)
```

# Data Importing and Cleaning

## Reading and Writing CSV Files

```{r}
write.csv(df, "data.csv", row.names=FALSE) # Write CSV
df_new <- read.csv("data.csv") # Read CSV
print(df_new)
```

## Handling Missing Values

```{r}
df_missing <- data.frame(
  Name = c("Alice", "Bob", NA),
  Age = c(25, NA, 35),
  Score = c(90, 85, NA)
)

# Check for missing values
print(is.na(df_missing))

# Remove missing values
df_clean <- na.omit(df_missing)
print(df_clean)

# Replace missing values with mean
df_missing$Age[is.na(df_missing$Age)] <- mean(df_missing$Age, na.rm=TRUE)
print(df_missing)









